\documentclass{article}
\usepackage{amsmath,amsthm,amsfonts,microtype,amssymb}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage{mathpazo}
\usepackage{tikz-cd}
\usepackage{geometry}
\usepackage{hyperref}


\title{Homework 01 \\ Fundamentals of Fundamental Groups}
\author{Algebraic Topology - Winter 2021}
\date{Due: \textbf{January 21, 2021, 11:59 pm}}

\begin{document}
\pagenumbering{gobble}
\maketitle

In the following questions, all maps are continuous and all homotopies are basepoint preserving.
\begin{enumerate}
    \item Show that \textit{path-homotopy} is an equivalence relation.
    \item Let $(X,x)$ be a pointed topological space.
    
    \begin{enumerate}
        \item  
        Let $c_x$ denote the constant path at $x$ and let $\gamma$ be a loop in $X$ based at $x$.
        Explicitly construct a reparametrization $\varphi: [0,1] \to [0,1]$ such that $\gamma \cdot c_x = \gamma \circ \varphi$.
        \item  
        Let $\gamma_1$, $\gamma_2$, $\gamma_3$ be loops in $X$ based at $x$.
        Explicitly construct a reparametrization $\varphi: [0,1] \to [0,1]$ such that $(\gamma_1 \cdot \gamma_2) \cdot \gamma_3 = (\gamma_1 \cdot (\gamma_2 \cdot \gamma_3)) \circ \varphi$.
        \item \emph{Justify} your answers by drawing pictures.
    \end{enumerate}
    
    
    \item 
        Let $X$ be a topological space and let $A$ be a subspace of $X$. Let $a_0$ be a point in $A$. We say that a map $r: X \to A$ is a \emph{retraction} if $r(a) = a$ for all $a \in A$.
        \begin{enumerate}
            \item What can you say about $\pi_1(r) : \pi_1(X, a_0) \to \pi_1(A, a_0)$ if $r$ is a retraction?
        \end{enumerate}
        We can post-compose $r$ with the inclusion of $A$ in $X$ to get a map $r' : X \to X$. We say that $r$ is a \emph{deformation retraction} if it is a retraction and $r'$ is homotopic to the identity map.
        \begin{enumerate}[resume]
            \item What can you say about $\pi_1(r) : \pi_1(X, a_0) \to \pi_1(A, a_0)$ if $r$ is a deformation retraction?
            \item 
            Come up with a concrete non-trivial example of a retraction which is not a deformation retraction, where non-trivial means that $A \neq \{ a_0 \}$. You do not have to provide proofs for this part.
        \end{enumerate}
        
        
    \item Show that homotopy equivalences satisfy the \emph{2-out-of-3 property}:\footnote{See also \url{https://ncatlab.org/nlab/show/category+with+weak+equivalences\#definition}.} given spaces $(X,x)$, $(Y,y)$, and $(Z,z)$ and maps $f:(X,x) \to (Y,y)$, and $g:(Y,y) \to (Z,z)$, if any two of $f$, $g$, and $g \circ f$ are homotopy equivalences, then so is the third.
    \begin{center}
        \begin{tikzcd}
            (X,x)   
            \arrow[r, "f", swap] 
            \arrow[rr, "g \circ f", bend left]
            &
            (Y,y)
            \arrow[r, "g",swap]  
            &
            (Z,z)
        \end{tikzcd}
    \end{center}
    % \item  Let $X$ be a topological space. Define the \emph{fundamental groupoid}, denoted $\Pi_1(X)$, to be the category whose objects are points in $X$ and whose morphisms are path-homotopy classes of paths. 
    % \begin{align*}
    %     \Pi_1(x,y) := \left\{ [\gamma] \mid \gamma: [0,1] \to X, \partial_0 \gamma = x, \partial_1 \gamma = y \right\}.
    % \end{align*}
    % \begin{enumerate}
    %     \item Recall that a groupoid is a category all of whose morphisms are invertible. 
    %     Verify that $\Pi_1(X)$ is a groupoid.
    % \end{enumerate}
    % Let $x$ be a point in $X$. We can think of the fundamental group, $\pi_1(X,x)$, as the category with exactly one object $x$ and with morphisms $\Pi_1(x,x)$.
    % \begin{enumerate}[resume]
    %     \item Show that if $X$ is path-connected, then the categories $\Pi_1(X)$ and $\pi_1(X,x)$ are equivalent.
    % \end{enumerate}
    % Thus, the fundamental groupoid is a basepoint-independent construction which is equivalent to the fundamental group. 
\end{enumerate}

\end{document}
